---
title: "03 Summarizing Data"
output: html_document
---
# Data Summary
## Summarizing Categorical data
- Relative frequency
	- Ratios of frequencies
	- Number of observations belonging to a category divided by total observations
	- Proportions, can be multiplied by 100 and expressed as percentage
- Distribution of a categorical variable
	- List of values with its associated count

Read in CSV into "fund" 

```{r}
fund <- read.csv("http://tiny.cc/retirment_fund")
```

See first six rows of data with **head**
```{r}
head(fund)
```
Display the **str**ucture of the dataset, see [Rdocumentation](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/str) for extra options and arguments
```{r}
str(fund)
```

*How many retirement funds are low, average, or high risk?*
Make a table of the Risk column in the Fund dataset, this will be a **frequency** table that shows the counts of each variable
```{r}
risk.tb <- table(fund$Risk)
```

Display new table risk.tb
```{r}
risk.tb
```

*What percentage of retirement funds are low, average, or high risk?*
Make a table to show the proportions, multiply by 100 to create percentage, round to two decimals
```{r}
risk.prop <- prop.table(risk.tb)*100
round(risk.prop,2)
```

View as barplot, use **main** to set title:
```{r}
barplot(risk.prop, main = "Level of Risk in Retirement Funds Sample")
```
Mode of Categorical Data
- Mode
  - Category with the highest frequency
  - Longest bar in the bar chart
  - Two or more categories can tie for the highest frequency
    - Bimodal (2) multimodal (2+)

## Summarizing Numerical Data
Summaries of numerical variables:
- Median
- Mean
- Lower quartile (25th percentile)
- Upper Quartile (75th percentile)
- Interquartile Range (IQR, 75th-25th percentile)

Use file.choose to have IDE open Finder and select file
```{r}
music <- read.csv(file.choose())
```

Use **Summary** to get some basic information about a variable, in this case Size..MB.
```{r}
summary(music$Size..MB.)
```

Notice large Max outlier, explains why Mean is larger than Median

Create a histogram
```{r}
hist(music$Size..MB.)
```
Let's look at standard deviation:
```{r}
sd(music$Size..MB.)
```
Find total size of music library by multiplying song count and size:
```{r}
500*mean(music$Size..MB.)
```

95% of data will be between +- 2 standard deviations (STD)

```{r}
500*(2*sd(music$Size..MB.) + mean(music$Size..MB.))
```

## Shape of a Distribution
- Symmetrical data
  - 50% of date below mean, 50% above mean
- Right skewed
  - Not symmetrical and more outliers to right
- Left skewed
  - Outliers to left
- Area under curve
  - 68% of data is between -1 to 1 STD
  - 95% of data is between -2 to 2 STDs
  - 99.7% of data is between -3 to 3 STDs