---
title: "04 Probability"
output: html_document
---
# 03 Probability 
## Introduction
- **Probability** is the measure of long run relative frequency 
  - In the long run, the accumulated relative frequency converges to a constant
- **Law of Large Numbers (LLN)**
  - The relative frequency of an outcome converges to a number
  - Only applies in the long run
  
```{r}
# Set seed that creates randomness
set.seed(123)

n = 10

# Sample is 0 and 1 (range between 0 and 1)
# n = sample size
# p assigns .3 probability to 0 and .7 to 1
# Sampling is with replacement, each is independent
x <- sample(0:1, n, p = c(0.3,0.7), replace = TRUE)

# Print x
x
```
  
Calculate cumulated relative frequency and feed it into sumx
  
```{r}
sumx <- cumsum(x)
```

c() creates empty vector

```{r}
cumMean <- c()

# Iterate through 1:n calculating means and adding to cumMean
# sumx is cumulative sum, dividing by i gives mean
for (i in 1:n) {
  cumMean[i] <- sumx[i]/i
}

# Plot it
plot(cumMean, type = "l")
```

There's a lot of variance, change n to 100 to see LLN in action:
```{r}
n = 100
```

Re-execute above code with new n
```{r}
plot(cumMean, type = "l")
```
  Now much less separation, we are arriving at a specific mean
  ###Rules for Probability
  - Sample space
    - Set of all possible outcomes
    - E.g. rolling a die, outcomes are 1:6
      - Coin flip is heads or tails
    - Denoted by S; e.g. S={heads, tails}
    - Subsets of sample spaces are **events**
      - Denoted as A, B, etc...
      - E.g. heads is event A, tails B
  - Venn Diagrams 
    - The probability of an event A is denoted P(A)
    - Venn diagrams are graphs for depicting the relationship among events
  - Addition rule for disjoint events
    - Disjoint events are mutually exclusive, no outcomes in common
    - The union of two events is the collection of outcomes in A, in B, or in
      both (A and B)
    - The venn diagrams don't touch
    - P(A or B) = P(A) + P(B)
    - Can extend to more than two events (E)
      - P(E1 or E2 or E3...) = P(E1) + P(E2)...
  - Complement Rule
    - The complement of event A consists of the outcomes in S but not in A 
    - Denoted as AC (superscript C)
    - P(A) = 1 - P(AC)
  - Addition Rule
    - For two events that are not disjoint A and B
    - The probability that one or the other occurs is the sum of the probability
      minus the probability of their intersection
    - Venn Diagram minus the middle 
    - P(A or B) = P(A) + P(B) - P(A and B)
  
  
  ## Conditional Probability
  - Contingency Table
    - Show counts and sums of two variables
  - Probability table
    - Divide cells in contingency tables by totals to determine probability
    - Replace all cell counts with probabilities to create joint probability 
  - Joint Probability
    - Represents the probability of an intersection of two or more events 
      - Combination of attributes
  - Marginal probablity 
    - The probability of observing an outcome with a single attribute regardless 
    of other attributes
    - In the joint probability table, this are the total cells
  - Conditional probabilities 
    - "The probability of A given B" is written as P(A | B)
    - P(A | B) = P(A and B)/P(B)
    - Example: Are purchases more likely than Comcast?
      - Yes = successful purchase
      - P(Yes | Comcast) = P(Yes and Comcast)/P(Comcast)
  - Independent Events
    - Two events are independent if the occurrence of one does not affect the
      chances for the occurrence of the other
    - Events that are not independent are dependent
    - Two events A and B are independent if the probability that both A and B
      occur is the product of the probabilities of the two events
      - P(A and B) = P(A) * P(B)
    - Example: Probability that a breakdown on an assembly line will occur in 
      the next five days, interfering with the completion of an order?
      - Past data indicates a 95% chance the assembly doesn't break in a day
      - Assuming days are independent:
        - P(OK for 5 days) = 95**5 = .774
      - Probability that it WILL occur is 1 - .774 = .226 (complement rule)
  - Dependent Events
    - Dependent 
      If A and B are dependent:
        P(A/B) = P(A and B) / P(B)
        P(B/A) = P(A and B) / P(A)
    - **Multiplication Rule**
      - P(A and B) = P(B | A) * P(A)
      - P(A and B) = P(A | B) * P(B)
    - Order in Conditional Probabilities 
      - Imagine a table of ads seen vs tv show
      - If the viewer sees ads, what is the chance they are watching Survivor?
        - Find P (Survivor | See Ads)
          = P(Survivor and See Ads)/P(See Ads)
  ### Bayes Theorem
  - P(Bi | A) = P(A | Bi)P(Bi)/P(A)
  - Bi = Different dimensions you're examining
    - Columns in a contingency table
    - Example: Probability of cancer given a test is positive
      - P(Cancer | Test positive)
      - P(Test negative | No cancer) = .925
      - P(Test positive | Cancer) = .85
      - P(Cancer) = .005
        |         |Positive | Negative|Total|
        |:--------|:--------|:--------|----:|
        |Cancer   |.00425   |         |.005 |
        |No Cancer|         |.92      |.995 |
        |Total    |         |         |1    | 
    - P(Cancer and positive) = P(Positive | Cancer)*P(Cancer)
    - The probability of positive given cancer * the probability of having
      cancer
       P(Positive | Cancer) =.85
       P(Cancer) = .005
       .85 * .005
      = .00425
    - P(No cancer and negative) = .925 * .995 = .92
    - With those two values you can fill out the rest of the contingency table
        |         |Positive | Negative|Total|
        |:--------|:--------|:--------|----:|
        |Cancer   |.00425   |.00075   |.005 |
        |No Cancer|.074625  |.92      |.995 |
        |Total    |.078875  |.921125  |1    | 
        
    - Calculate marginal probabilities by dividing inner cells by totals
      - P(Cancer | Positive) = P(Cancer and Positive) / P(Positive)
      - P(Positive) is the total of the positive column
      - The probability of positive AND cancer / the probability of positive
      - .00425/.078875 = .053
        